---
title: "Probelm 14-9 and 11"
author: "Chenghan Sun"
date: "3/15/2019"
output: html_document
---

# 14-9
## load data 
```{r}
setwd("/Users/cs0823/Documents/STA_207/dataset")
data = read.table(file = "Data14_9.csv", header= FALSE, sep = ",")
colnames(data)[colnames(data)=="V1"] <- "Ability"
colnames(data)[colnames(data)=="V2"] <- "Stability"
```

## (a) Find the maximum likelihold estimates of beta_0 and beta_1
```{r}
model=glm(Ability~Stability, family="binomial", data= data)
coef(model)
```
> from the results above: b0 = -10.30892518 b1 = 0.01891983
  Fitted response function: 
  
$$ \pi_{i} = \frac{exp(XB) }{1+ exp(XB))} = [1+exp(-XB)]^{-1} \ where \ XB = -10.31 +0.019X_{i} $$ 

## (b) Obtain a scater plot of the data 
```{r}
n=nrow(data)
Beta = as.matrix(coef(model), ncol=1, nrow=2)
X = cbind( rep(1,n), data$Stability )
pi_fitted = ( 1 + exp(-1* (X%*%Beta) ) )^(-1)
plot(y = pi_fitted, x=data$Stability)
lines(lowess(data$Stability, data$Ability), col=2)
```

## (c) Obtain exp(b1) and interpret this number
```{r}
exp(coef(model)[2])
```
> The odds of a stability are estimated to increase by 1.91% with one stability score increases.

## (d)
```{r}
x = as.vector(c(1, 550))
pi_fitted = (1 + exp(-1*t(x)%*%Beta))^(-1)
pi_fitted 
```
> The estimated probability to complete the task at Stability score =550 is 0.524.

## (e)  Estimate the emotional stability test score for which 70 percent of the employees with this test score are expected to be able to perform in a task group
```{r}
logit = function(p) { p / (1-p) }
b0 =coef(model)[1]
b1 =coef(model)[2]
score = (logit(0.7) - b0 ) / b1
score
```
> The stability score to have probability of 70% of employees to complete the tasks is estimated at 668.2015.

# 14-11
## load data
```{r}
setwd("/Users/cs0823/Documents/STA_207/dataset")
data = read.table(file = "Data14_11.csv", header= FALSE, sep = ",")
colnames(data)[colnames(data)=="V1"] <- "X"
colnames(data)[colnames(data)=="V2"] <- "Number"
colnames(data)[colnames(data)=="V3"] <- "Y"
```

## (a) Plot the estimated proportions Pj = Y.j / nj 
```{r}
P = data$Y/500
data$P = P
plot(data$X, data$P)
```
> The logistic response function seems to be appropriate.

## (b) Find the maximum likelihood estimates of beta_0 and beta_1
```{r}
model = glm(data$P~data$X, family=binomial(link="logit"), data=data)
coef(model)
```
> from the results above: 
Model: 
$$ \hat{\pi} = \frac{\exp(-2.0766+0.1359X_i)}{1+\exp(-2.0766+0.1359X_i)} $$

## (c) Obtain a scatter plot of the data with the estimated proportions from part (a)
```{r}
data_new = data.frame(X = seq(min(data$X), max(data$X),len=6))
data_new$P = predict(model, data_new, type="response")

plot(data_new$X, data_new$P)
lines(P ~ X, data_new)
```
> the logistic response function seems to fit well.

## (d) Obtain exp(b1) and interpret this number
```{r}
exp(coef(model)[2])
```
> The odds that a bottle is returned increases by about 14.5% with each additional 1 cent increase in deposit level.

## (e) What is the estimated probability that a bottle will be returned when the deposit is 15 cents?
```{r}
b0 = -2.0765647
b1 = 0.1358508 
exp(b0+b1*15) / (1 + exp(b0+b1*15))
```
> the estimated probability is 0.4903005.

## (f) Estimate the amount of deposit for which 75 percent of the bottles are expected to be returned
```{r}
(log(0.75 / (1 - 0.75)) - b0) / b1
```
> the amount of deposit is estimated to be 23.37253 ~ 23.  

