---
title: "Nested_Designs"
author: "Chenghan Sun"
date: "2/28/2019"
output: html_document
---

##26-4 

### Data Loading 
```{r}
library(Matrix)
library(lme4)
library(ALSM)
library(SuppDists)
library(ggplot2)
setwd("/Users/cs0823/Documents/STA_207/dataset")
data = read.table(file = "Data26_4.csv", header= FALSE, sep = ",")
colnames(data)[colnames(data)=="V1"] <- "Cases"
colnames(data)[colnames(data)=="V2"] <- "Machine"
colnames(data)[colnames(data)=="V3"] <- "Operator"
colnames(data)[colnames(data)=="V4"] <- "Days"
```

###a 
Obtain the residuals for nested design model with fixed factor effects and plot them against the fitted values.
###b 
Prepare alig~cd residual dot plots by machince.
```{r}
data$Machine = factor(data$Machine)
data$Operator= factor(data$Operator)

modelfix = aov(Cases ~ Machine + Machine/Operator, data=data)
summary(modelfix)

modelfix$residuals
qqnorm(modelfix$residuals)
qqline(modelfix$residuals)

aligned.dot.plot2(modelfix$residuals,data$Machine)
```
The nested model seems to be appropriate. And the aligned plot supports the conclusion. Normality assumption does not hold. The variance of residual seems to be constant across the groups.

#26-5

###a 
Can the operator effects be distinguished from the effects of shift~ in this study?
No, operator effects can not be distinguisted from the effect of shift.

###b
Plot the data
```{r}
ggplot(data = data, aes(x= Cases, y= Machine, colour=Operator))+
  geom_point() +  geom_jitter() 
```
> From the figure above, we can see that "Machine 3" can produce more cases. 

###c
Obtain the analysis of variance table
```{r}
summary(modelfix)
```

###d
Test whether or not the mean outputs differ for the three machine models.
```{r}
# from the anova table above
MSA = 847.8
MSE = 23.6
F_test = MSA/MSE
F_test

df_SSA = 2
df_SSE = 48
p_value = pf(F_test, df1 = df_SSA, df2 = df_SSE, lower.tail = F)
p_value
```
we want to test the following hypothesis:
$$ Ho : \sigma^2_{\alpha} = 0  $$
$$ Ha : \sigma^2_{\alpha} \neq  0  $$
For fixed B nested in fixed A, the F* value is calculated to be 35.9. 
Since the p value of this test is 2.9e-10 smaller than 0.01, so we reject the Ho and support Ha where there is variation in mean outputs among three machines.

###e
Test whether or not the mean outputs differ for the operators assigned to each machine
```{r}
MSAB = 252.5
F_test = MSAB/MSE
F_test 

df_SSAB = 9
p_value = pf(F_test, df1 = df_SSAB , df2 = df_SSE, lower.tail = F)
p_value
```
we want to test the following hypothesis:

$$ Ho : \sigma^2_{B(A)} = 0  $$
$$ Ha : \sigma^2_{B(A)} \neq  0  $$
For fixed B nested in fixed A, F* is calculated to be 10.70. Since the p value of this test is 6.98e-10 and smaller than 0.01, we reject the Ho and support Ha where there is variation in mean outputs among operators.

###f
Test for each machine separately whether or not the mean outputs for the four operators differ.
```{r}
df= data.frame(fitted_values = modelfix$fitted.values)
df$Machine = data$Machine
df$Operator = data$Operator
Ai_means = aggregate(Cases~Machine, FUN= mean,  data = data )

df_new = merge(x = df, y=Ai_means, by = "Machine", all_y = T)
colnames(df_new)[4] = c("Ai_means")
df_new$B_Ai = df_new$fitted_values - df_new$Ai_means

# decomposition of SSB(Ai)
SSB_Ai = aggregate(B_Ai ~ Machine, data = df_new, FUN = function(x) { 
  sum(x^2) 
  } 
)

d=4
SSB_Ai$df = c(d-1)
SSB_Ai$MS =  SSB_Ai$B_Ai / SSB_Ai$df
SSB_Ai$F_test = SSB_Ai$MS / MSE  

SSB_Ai$p_value = pf(SSB_Ai$F_test, df1 = SSB_Ai$df, df2 = df_SSE, lower.tail = F)
SSB_Ai
```
we want to test the following:

(1)
$$ Ho : \sigma^2_{\beta(A1)} = 0 \  v.s  \ Ha : \sigma^2_{\beta(A1)} \neq  0  $$
we find p_value = 1.285e-04, which is smaller than 0.01, so reject H0 and conclude that there is operator effect on machine 1.

(2)

$$ Ho : \sigma^2_{\beta(A2)} = 0 \  v.s  \ Ha : \sigma^2_{\beta(A2)} \neq  0  $$
we find p_value = 4.941e-09, which is smaller than 0.01, so reject H0 and conclude that there is operator effect on machine 2.

(3)

$$ Ho : \sigma^2_{\beta(A3)} = 0 \  v.s  \ Ha : \sigma^2_{\beta(A3)} \neq  0  $$
we find p_value = 0.142, which is larger than 0.01, so reject Ha and conclude that there is no operator effect on machine 3.

###g
What is the family level of significance for the combined tests in parts (d). (e), and (f) using the Bonferroni inequality.

There will be 5 tests from parts (def), so by using Bonferroni procedure, and family level of significance should be smaller than 5*0.01 = 0.05.

