---
title: "Random_Mixed_Model"
author: "Chenghan Sun"
date: "2/24/2019"
output: html_document
---
# 25-16
Data-Loading
```{r}
setwd("/Users/cs0823/Documents/STA_207/dataset")
data = read.table(file = "Data25_16.csv", header= FALSE, sep = ",")
colnames(data)[colnames(data)=="V1"] <- "Time"
colnames(data)[colnames(data)=="V2"] <- "Technician"
colnames(data)[colnames(data)=="V3"] <- "Make_of_device"
colnames(data)[colnames(data)=="V4"] <- "Rep"
```

(a) Test whether or not the two factors interact
```{r}
a = 3
b = 3
rep = 5

fitted_values_generator=function(data) {
  fitted_values = c()
  for (i in 1:a) {
    for (j in 1:b) {
      data.subset = data[data[,2]==i & data[,3]==j,]
      fitted_values = c(fitted_values, rep(mean(data.subset[,1]), rep))
    }
  }
  return(fitted_values)
}

data$fitted = fitted_values_generator(data)
data$res = data$Time - data$fitted

#factor effects 
overall_mean = mean(data$Time)
A_means = tapply(data$Time, data$Technician, mean)
B_means = tapply(data$Time, data$Make_of_device, mean)
A_effects = A_means - overall_mean
B_effects = B_means - overall_mean

#interaction effects 
AB_effects = tapply(data$Time,list(data$Technician, data$Make_of_device), mean)
for (i in 1:a) {
  for (j in 1:b) {
    AB_effects[i,j] = AB_effects[i,j] - (overall_mean + A_effects[i] + B_effects[j])
  }
}

# MSE
n = 5
MSE = sum(data$res^2) / ((n-1)*a*b)
MSE

# MSAB
SSAB = 0
for (i in 1:a) {
  for (j in 1:b) {
    SSAB = SSAB + n * AB_effects[i,j]^2
  }
}

MSAB = SSAB / ((a-1)*(b-1))
MSAB

# F-statistic
F_test = MSAB/MSE
F_test

# Critical value
alpha = 0.01
qf(1-alpha, (a-1)*(b-1), (n-1)*a*b)

# p-value
pf(MSAB/MSE, (a-1)*(b-1), (n-1)*a*b, lower.tail=FALSE)
```
 > H0: sigma^2_alpha_beta = 0 vs. Ha: sigma^2_alpha_beta > 0. Since F* is calculated to be 5.84, which is larger than critical value of 3.89 for alpha = 0.01, reject H0 and conclude that there is indeed an interaction effect. The p-value is approximately 0.000994.

(b) find point estimate of $\sigma^2_{\alpha\beta}$
A point estimate of $\sigma^2_{\alpha\beta}$ is $s^2_{\alpha\beta}=\frac{MSAB-MSE}{n}$

```{r}
(MSAB - MSE)/n
```
 
 (c) Test whether or not factor A main effects are present
```{r}
MSA = n * b * sum(A_effects^2) / (a-1)
MSA

# F-statistic
F_test = MSA/MSE
F_test

# Critical value
alpha = 0.01
qf(1-alpha, a-1, (n-1)*a*b)
```
 > H0: sigma^2_alpha = 0 vs. Ha: sigma^2_alpha > 0. Since F* is calculated to be 0.236, which is smaller than critical value of 5.248 for alpha = 0.01. We fail to reject the null hypothesis. It is important to test for the presence of factor A main effects since factor A is random and, if its effects are present, we would need a mixed model since factor B is fixed.
 
(d) Test whether or not factor B main effects are present
```{r}
MSB = n * a * sum(B_effects^2) / (b-1)
MSB

# F-statistic
F_test = MSB/MSAB
F_test

# Critical value
alpha = 0.01
qf(1-alpha, b-1, (a-1)*(b-1))
```
> H0: beta_j = 0 for j = 1,2,3 vs. Ha: not all beta_j = 0. Since F* is calculated to be 0.047, which is smaller than critical value of 18 for alpha = 0.01. We fail to reject the null hypothesis. It is important to test for the present of factor B main effects since factor B is fixed and, if its effects are present, we would need a mixed model since factor A is random.

(e) Use the Tukey procedure and a 95 percent family confidence coefficient to make all pairwise comparisons. 
```{r}
# Pairwise differences between means
D1 = B_means[1] - B_means[2]
D2 = B_means[1] - B_means[3]
D3 = B_means[2] - B_means[3]
c(D1,D2,D3)

# Standard error of the pairwise differences
stand_err = sqrt(2 * MSAB / (a * n))
stand_err

# Multiplier
alpha = 0.05
T_test = qtukey(1-alpha, b, (a-1)*(b-1)) / sqrt(2)
T_test

# Confidence intervals
c(D1 - T_test*stand_err, D1 + T_test*stand_err)
c(D2 - T_test*stand_err, D2 + T_test*stand_err)
c(D3 - T_test*stand_err, D3 + T_test*stand_err)
```
> We are 95% confidence that the true pairwise differences between the means for the three disk drive were calcualted above. Since each of the confidence intervals for the pairwise comparisons between the means for the three disk drive makes contains 0 at the 0.05 level of significance, we can conclude there is no difference between drive 1 and drive 2, drive 1 and 3, and drive 2 and 3.

(f) Use the Satterthwaite procedure to obtain an approximate 99 percent confidence interval for mu_dot_1
```{r}
# Marginal mean
mu_1 = B_means[1]

# Standard error of marginal mean 
c1 = (a-1) / (n*a*b)
c2 = 1 / (n*a*b)
stand_err_mu_1 = sqrt(c1*MSAB + c2 * MSA)
stand_err_mu_1

# degrees of freedom
df = (c1*MSAB + c2*MSA)^2 / ((c1*MSAB)^2/((a-1)*(b-1)) + (c2*MSA)^2/(b-1))
df

# Multiplier
alpha = 0.01
T_test = qt(1-alpha/2, df)
T_test

# Confidence intervals
c(mu_1 - T_test*stand_err_mu_1, mu_1 + T_test*stand_err_mu_1)
```
> From the calculation above, we are 99% confidence that the marginal mean for disk 1 using the Satterthwaite procedure is between 39.485 and 72.781. 












