---
title: "Three_Factor"
author: "Chenghan Sun"
output: html_document
---
> Data loading 

```{r}
setwd("/Users/cs0823/Documents/STA_207/HW4")
data = read.table(file = "Data24_12.csv", header= FALSE, sep = ",")
colnames(data)[colnames(data)=="V1"] <- "Y"
colnames(data)[colnames(data)=="V2"] <- "A"
colnames(data)[colnames(data)=="V3"] <- "B"
colnames(data)[colnames(data)=="V4"] <- "C"
```

# 24-12
part (a)
```{r}
X_A_fac = factor(data$A)
X_B_fac = factor(data$B)
X_C_fac = factor(data$C)
Y = data$Y
fit = aov(Y ~ X_A_fac+X_B_fac+X_C_fac+X_A_fac*X_B_fac+X_A_fac*X_C_fac+X_B_fac*X_C_fac+X_A_fac*X_B_fac*X_C_fac, data=data)
values = fitted.values(fit)
res = fit$residuals
```

> The residuals based on ANOVA model 24-14:
> res
    1     2     3     4     5     6     7     8     9    10    11    12    13    14    15    16    17    18    19    20 
 31.4 -43.6  17.4  20.4 -25.6 -30.0  48.0  18.0 -55.0  19.0  44.8 -23.2 -33.2  20.8  -9.2  -3.4 -12.4   0.6 -25.4  40.6 
   21    22    23    24    25    26    27    28    29    30    31    32    33    34    35    36    37    38    39    40 
 -1.2 -28.2 -17.2  13.8  32.8 -18.2  15.8   5.8  25.8 -29.2  29.6  39.6 -32.4 -34.4  -2.4  -6.6 -22.6  10.4  21.4  -2.6 
   41    42    43    44    45    46    47    48    49    50    51    52    53    54    55    56    57    58    59    60 
 27.6 -34.4 -26.4  50.6 -17.4  -4.6  12.4  25.4 -34.6   1.4   0.6  -0.4  14.6 -20.4   5.6 -19.4   4.6 -43.4  50.6   7.6 

part(b) Prepare normal probability plot for the residuals 
```{r}
quantiles = qnorm(((1:length(res)) - 0.5) / length(res))
plot(quantiles,sort((res - mean(res)) / sd(res)), xlab="Theoretical Quantiles", ylab="Sample Quantiles", main = "Q-Q Plot")
abline(0,1)

cor(quantiles,sort(res))
```
> The coefficient of correlation is found to be 0.9906479.
> The residuals do not seem to deviate much from the theoretical line on the normal Q-Q plot. Thus, The normality assumption is valid. 

24-13
part (a) AB plots 
```{r}
library(dae)
avfactorA = tapply(Y, X_A_fac, mean)
avfactorB = tapply(Y, X_B_fac, mean)
avfactorC = tapply(Y, X_C_fac, mean)
# define levels of factor A&B&C, a&b&c, and sample size n
a = 2
b = 3
c = 2
n = 5

Alevel = 1:a
Blevel = 1:b
Clevel = 1:c
par(mfrow=c(3,3))
# main effec plot of factor A
plot(Alevel, avfactorA, type="o", xlab = "factor A", ylab = "Y", main= "Main Effect of factor A")
# main effec plot of factor B
plot(Blevel, avfactorB, type="o", xlab = "factor B", ylab = "Y", main= "Main Effect of factor B")
# main effec plot of factor C
plot(Clevel, avfactorC, type="o", xlab = "factor C", ylab = "Y", main= "Main Effect of factor C")
# two-way interaction plots
interaction.plot(X_A_fac, X_B_fac, Y, xlab = "factor A", ylab = "Y", trace.label = "factor B", 
                 main= "Interaction Plot of factor A and B")
interaction.plot(X_A_fac, X_C_fac, Y, xlab = "factor A", ylab = "Y", trace.label = "factor C", 
                 main= "Interaction Plot of factor A and C")
interaction.plot(X_B_fac, X_C_fac, Y, xlab = "factor B", ylab = "Y", trace.label = "factor C", 
                 main= "Interaction Plot of factor B and C")

# three-way interaction plots
#interaction.ABC.plot(Y, X_A_fac, X_B_fac, X_C_fac, data = data, title="A:B:C Interaction Plot")


```
> We can conclude all the interaction terms are presented. 

part (b)
```{r}
anova(fit)
```

part (c)
```{r}
X_A = data$A
X_B = data$B
X_C = data$C 
rep = data$V5
data_fitted = data.frame(values, X_A, X_B, X_C, rep)
#uhat_ijk = data_fitted$values

uhat_ddd = 0
for (i in 1:a){
  for (j in 1:b){
    for (k in 1:c){ 
      for (m in 1:n){
        uhat_ddd = uhat_ddd+(data_fitted[which(data_fitted$X_A==i & data_fitted$X_B==j 
                                               & data_fitted$X_C==k & data_fitted$rep==m),]$values)
      }
    }
  }
}
uhat_ddd = uhat_ddd/n/a/b/c

uhat_idd = 0
for (j in 1:b){
  for (k in 1:c){ 
    for (m in 1:n){
      uhat_idd = uhat_idd+(data_fitted[which(data_fitted$X_B==j 
                                             & data_fitted$X_C==k & data_fitted$rep==m),]$values)
    }
  }
}
uhat_idd = uhat_idd/n/b/c
alpha_i = uhat_idd - uhat_ddd
```

```{r}
uhat_djd = 0
for (i in 1:a){
  for (k in 1:c){ 
    for (m in 1:n){
      uhat_djd = uhat_djd+(data_fitted[which(data_fitted$X_A==i 
                                             & data_fitted$X_C==k & data_fitted$rep==m),]$values)
    }
  }
}
uhat_djd = uhat_djd/n/a/c
beta_j = uhat_djd - uhat_ddd
```

```{r}
uhat_ddk = 0
for (i in 1:a){
  for (j in 1:b){ 
    for (m in 1:n){
      uhat_ddk = uhat_ddk+(data_fitted[which(data_fitted$X_A==i 
                                             & data_fitted$X_B==j & data_fitted$rep==m),]$values)
    }
  }
}
uhat_ddk = uhat_ddk/n/a/b
gamma_k = uhat_ddk - uhat_ddd
```

```{r}
uhat_ijd = 0
for (k in 1:c){ 
  for (m in 1:n){
    uhat_ijd = uhat_ijd+(data_fitted[which(data_fitted$X_C==k & data_fitted$rep==m),]$values)
  }
}
uhat_ijd = uhat_ijd/n/c
uhat_ijd_new = matrix(1, nrow = 2, ncol = 3)
uhat_ijd_new[1,] = c(uhat_ijd[1], uhat_ijd[2], uhat_ijd[3])
uhat_ijd_new[2,] = c(uhat_ijd[4], uhat_ijd[5], uhat_ijd[6])

alpha_beta_ij = matrix(1, nrow = 2, ncol = 3)
for (i in 1:a){
  for (j in 1:b){
    alpha_beta_ij[i,j] = uhat_ijd_new[i,j] - uhat_idd[i] - t(uhat_djd)[j] + uhat_ddd
  }
}
```

```{r}
uhat_idk = 0
for (j in 1:b){ 
  for (m in 1:n){
    uhat_idk = uhat_idk+(data_fitted[which(data_fitted$X_B==j & data_fitted$rep==m),]$values)
  }
}
uhat_idk = uhat_idk/n/b
uhat_idk_new = matrix(1, nrow = 2, ncol = 2)
uhat_idk_new[1,] = c(uhat_idk[1], uhat_idk[2])
uhat_idk_new[2,] = c(uhat_idk[3], uhat_idk[4])

alpha_gamma_ik = matrix(1, nrow = 2, ncol = 2)
for (i in 1:a){
  for (k in 1:c){
    alpha_gamma_ik[i,k] = uhat_idk_new[i,k] - uhat_idd[i] - t(uhat_ddk)[k] + uhat_ddd
  }
}
```


```{r}
uhat_djk = 0
for (i in 1:a){ 
  for (m in 1:n){
    uhat_djk = uhat_djk+(data_fitted[which(data_fitted$X_A==i & data_fitted$rep==m),]$values)
  }
}
uhat_djk = uhat_djk/n/a
uhat_djk_new = matrix(1, nrow = 3, ncol = 2)
uhat_djk_new[1,] = c(uhat_djk[1], uhat_djk[2])
uhat_djk_new[2,] = c(uhat_djk[3], uhat_djk[4])
uhat_djk_new[3,] = c(uhat_djk[5], uhat_djk[6])

beta_gamma_jk = matrix(1, nrow = 3, ncol = 2)
for (j in 1:b){
  for (k in 1:c){
    beta_gamma_jk[j,k] = uhat_djk_new[j, k] - uhat_djd[j] - t(uhat_ddk)[k] + uhat_ddd
  }
}
```

> make the dimension of all matrices same 

```{r}
# initilize squences
# from the previous results 
sq_uhat_ijk = c(1218.6, 1051.0, 1274.2, 1122.4, 1218.2, 1051.2, 1036.4, 870.6, 1077.4, 931.6, 1020.4, 860.4)
sq_uhat_ddd = c(rep(1061.033, 12))

sq_alpha_i = c(rep(alpha_i[1], 6), rep(alpha_i[2], 6))
sq_beta_j = c(rep(c(-16.88333, -16.88333, 40.36667, 40.36667, -23.48333, -23.48333),2))
sq_gamma_k = c(rep(c(79.83333, -79.83333), 6))

sq_alpha_beta_ij = c(-4.25, -4.25, 2, 2, 2.25, 2.25, 4.25, 4.25, -2, -2, -2.25, -2.25)
sq_alpha_gamma_ik = c(rep(c(1.233333, -1.233333),3), rep(c(-1.233333, 1.233333),3))
sq_beta_gamma_jk = c(rep(c(3.516667, -3.516667, -5.433333, 5.433333, 1.916667, -1.916667),2))

alpha_beta_gamma_ijk = sq_uhat_ijk - (sq_uhat_ddd + sq_alpha_i+ sq_beta_j+ sq_gamma_k+ 
                                        sq_alpha_beta_ij+ sq_alpha_gamma_ik + sq_beta_gamma_jk)
```

```{r}
abg_ijk_data = data.frame(alpha_beta_gamma_ijk, c(rep(1,6), rep(2,6)), c(1,1,2,2,3,3,1,1,2,2,3,3),
                          c(1,2,1,2,1,2,1,2,1,2,1,2))

colnames(abg_ijk_data)[colnames(abg_ijk_data)=="c.rep.1..6...rep.2..6.."] <- "index_i"
colnames(abg_ijk_data)[colnames(abg_ijk_data)=="c.1..1..2..2..3..3..1..1..2..2..3..3."] <- "index_j"
colnames(abg_ijk_data)[colnames(abg_ijk_data)=="c.1..2..1..2..1..2..1..2..1..2..1..2."] <- "index_ik"

SSABC = 0
for (i in 1:a){
  for (j in 1:b){
    for (k in 1:c){
      SSABC = SSABC + (abg_ijk_data[i,j,k])^2
    }
  }
}

MSABC = 2*SSABC/(a-1)/(b-1)/(c-1)
```

> H0: all (alpha_beta_gamma)ijk equal 0; Ha: not all (alpha_beta_gamma)ijk equal 0.\
> From the analysis above, MSABC = 9.517, MSE = 858.04, so F_star = MSABC/MSE = 0.01.\
F(0.95; 2, 48) = 3.19. Since F_star << F(0.95; 2, 48), conclude H0. The p_value = 0.99.\

24.13
part (d) test AB,AB, BC interactions
From the previous ANOVA table and calculation,
1. we have $H_0: (\alpha\beta)_{ij}=0,\forall i,j$ vs. $H_1$: not all $(\alpha\beta)_{ij}=0$,
$F^* = MSAB/MSE = 0.32$,
Reject $H_0$ if $F^*>F(0.95,2,60-12)=3.19$,
So we cannot reject $H_0$. $P = 0.7$.

2. we have $H_0: (\alpha\gamma)_{ij}=0,\forall i,j$ vs. $H_1$: not all $(\alpha\hamma)_{ij}=0$,
$F^* = MSAC/MSE = 0.11$,
Reject $H_0$ if $F^*>F(0.95,1,60-12)=4.04$,
So we cannot reject $H_0$. $P = 0.8$.

3.we have $H_0: (\abeta\gamma)_{ij}=0,\forall i,j$ vs. $H_1$: not all $(\beta\hamma)_{ij}=0$,
$F^* = MSBC/MSE = 0.53$,
Reject $H_0$ if $F^*>F(0.95,2,60-12)=3.19$,
So we cannot reject $H_0$. $P = 0.6$.


part (e) test A,B,C main effects.
1. $H_0: \alpha_i=0,\forall i$ vs. $H_1$: not all $\alpha_i=0$,
$F^* = MSA/MSE = 629.76$,
Reject $H_0$ if $F^*>F(0.95,1,60-12)=4.04$,
Therefore, we reject $H_0$, $P < 0.001$.

2. $H_0: \beta_j=0,\forall j$ vs. $H_1$: not all $\beta_j=0$.
$F^* = MSB/MSE = 28.74$,
Reject $H_0$ if $F^*>F(0.95,2,60-12)=3.19$,
Therefore, we reject $H_0$, $P < 0.001$.

3. $H_0: \gamma_k=0,\forall k$ vs. $H_1$: not all $\gamma_k=0=0$,
$F^* = MSC/MSE = 445.67$
Reject $H_0$ if $F^*>F(0.95,1,60-12)=4.04$
Therefore, we reject $H_0$, $P < 0.001$

part (f)
There seems not to exist ABC, AB, AC,BC interaction term, and main effects $\alpha$, $\beta$ seem to exist.
```{r}
1-(1-0.05)^7
```
The upper bount is 0.30167

part (g)
The result from f confirms our analysis in part a.

24.14
part (a)
```{r}
#calculate Y bar
mean1 = aggregate(data$Y,list(data$A),mean)
mean2 = aggregate(data$Y,list(data$B),mean)
mean3 = aggregate(data$Y,list(data$C),mean)
# mean1
# mean2
# mean3

#calculate D estimate
D = 1:5
D[1] = 1155.9333-  966.1333
D[2] =  1044.15 - 1101.40
D[3] =  1044.15 - 1037.55
D[4] =  1101.40 - 1037.55
D[5] = 1140.867 - 981.200
cat("Estimate of D is",D,"\n")

#calculate muhat ijk.
muijk = c(1:12)
muijk = as.numeric(muijk)
for (i in (1:12)){
  yijk = sum(data$Y[((i-1)*5+1):(i*5)])
  muijk[i] = yijk/5
}

#calculate SSE
s = c(1:60)
for (i in c(1:12)){
  for (j in c(1:5)){
  s[(i-1)*5+j] = (data$Y[(i-1)*5+j] - muijk[i])^2
  }
}
SSE = sum(s)
MSE = SSE/(4*2*3*2)

#calculate s(D)
SD = 1:5
a=2
b=3
c=2
SD[1] = sqrt((2/(5*b*c))*MSE)
SD[2] = sqrt((2/(5*a*c))*MSE)
SD[3] = sqrt((2/(5*b*c))*MSE)
SD[4] = sqrt((2/(5*b*c))*MSE)
SD[5] = sqrt((2/(5*b*a))*MSE)
B = qt(1-0.05/10,48)

#construct C.I.
CIL = 1:5
CIU = 1:5
for(i in 1:5){
  CIL[i] = D[i] - B*SD[i]
  CIU[i] = D[i] + B*SD[i]
  cat("Confidence Interval of D",i,"is(",CIL[i],",",CIU[i],")\n")
}
```
We have $\bar{Y}_1... = 1155.9333$, $\bar{Y}_2... = 966.1333$, 
$\bar{Y}_.1.. = 1044.15$, $\bar{Y}_.2.. = 1101.40$ , $\bar{Y}_.3.. = 1037.55$,
$\bar{Y}_..1. = 1040.867$ , $\bar{Y}_..2. = 981.200$.

From the C.I., we can see that only D3 includes0. Thus, factor A means are different, factor C means are different. Factor B mean at level 2 is different from means at level 1 and 3.

part (b)
```{r}
means231 = aggregate(data$Y,list(data$A,data$B,data$C),mean)
hatmu231 = 1020.4
cat("muhat231 is",hatmu231,"\n")
t = qt(0.974,48)
cat("t value is",t,"\n")
s = sqrt(MSE/5)
cat("S(mu231) is",s,"\n")
CIL = hatmu231 - t*s
CIU = hatmu231 + t*s
cat("Confidence Interval is (",CIL,",",CIU,")\n")
```

